---
title: "R_raster"
author: "Jeff Grayum"
date: "3/24/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Link to example: https://www.neonscience.org/resources/learning-hub/tutorials/raster-data-r


Installing and loading libraries.
```{r}
#install.packages("raster")
#install.packages("sp")
#install.packages("rgdal")
library(raster)
library(sp)
library(rgdal)
library(tidyverse)
```


Loading dataset.
```{r}
#These remote sensing data files provide information on the vegetation at the National Ecological Observatory Network's San Joaquin Experimental Range and Soaproot Saddle field sites.

DEM <- raster("NEON-DS-Field-Site-Spatial-Data/SJER/DigitalTerrainModel/SJER2013_DTM.tif")

DEM

#dimensions: "size" of the file in pixels (nrows, ncols, total cells).
#resolution: size of each pixel (1 meter x 1 meter, in this example).
#extent: spatial extent of raster (same coordinate units as coordinate ref system of the raster).
#crs: coordinate reference system for the raster (UTM's, in this case).
```

Some notes:
-Raster data (or "gridded data") are data that are SAVED IN PIXELS.
-Each pixel represents an area on the Earth's surface.
-sp and raster packages are both needed.
-rgdal allows us to export rasters to GeoTIFF format.

```{r}
#This raster doesn't have min/max values associated with its attributes.  We must add them manually.

DEM <- setMinMax(DEM)

#Notice "values" is now listed as an attribute, and has min, max --> what these represent depends on what is represented by each pixel in the raster.

DEM
```

Let's view the rasters min and max values, and the range.
```{r}
cellStats(DEM, min)

cellStats(DEM, max)

cellStats(DEM, range)

#Can we do this tidyverse-style?
DEM %>%
  cellStats(min)

DEM %>%
  cellStats(max)

DEM %>%
  cellStats(range)

#We can!!
```

Now we must consider the coordinate reference system (crs), and the extent (exact boundries).
```{r}
DEM@crs

#The raster is located in UTM Zone 11

DEM@extent
```


